---
title: "Visualizations"
author: "Garrett Marshall"
date: "`r Sys.Date()`"
output: word_document
---


```{r message=FALSE}
library(tidyverse)
library(cowplot)
library(nflplotR)
library(forcats)
```


```{r}
data <- read.csv("/Users/garrett/Desktop/GitHub/NFLPlayPredictionProject/raw-data/games_pbp.csv")
view(data)
#str(data)
```

```{r}
data <- data %>%
  select( -c('X'))
```



```{r}
data1 <- data %>%
  group_by(game) %>%
  summarise(count= n()) %>%
  filter(grepl('bears_at_vikings_2022_reg_5', game))

```



```{r}
d <- data %>%
  group_by(possession_abbr, run_pass) %>%
  summarise(count=n())
```

```{r}
team_colors <- nflreadr::load_teams()
team_colors <- team_colors %>%
  select( c("team_abbr", "team_color") )

d <- inner_join(d, team_colors, by= c("possession_abbr"="team_abbr"))
```


```{r}
data_pass <- data %>%
  group_by(possession_abbr) %>%
  filter(run_pass == "pass") %>%
  summarise(pass_count= n())

data_run <- data %>%
  group_by(possession_abbr) %>%
  filter(run_pass == "run") %>%
  summarise(run_count= n())

run_pass_df <- inner_join(data_pass, data_run, by="possession_abbr")
```

```{r}
run_pass_df <- run_pass_df %>%
  mutate(num_pass_per_game = as.double(round(pass_count/16, 2) )) %>%
  mutate(num_run_per_game = as.double(round(run_count/16, 2) ) )
```

```{r}
d <- d %>% 
  mutate(bar_color = ifelse(run_pass == "run", team_color, "grey"))
```

```{r}
d_test <- d %>% 
  group_by( possession_abbr ) %>%
  mutate( total = sum(count) ) %>%
  ungroup()


d_test <- inner_join(d_test, data_run, by="possession_abbr")

d_test <- d_test %>% 
  mutate(run_prop = run_count / total) %>%
  select( c('possession_abbr', 'run_prop') )

#d <- inner_join(d, d_test, by="possession_abbr")
#d
```


```{r}
d$run_prop <- d_test$run_prop
```

```{r}
d
```


```{r fig.height=5, fig.width=8}
stack <- ggplot(d, aes(x = reorder(possession_abbr, -run_prop), y = count, fill = forcats::fct_rev(bar_color))) +
  geom_col(position = "fill") +
  scale_fill_identity(guide = "none") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Proportion of Pass and Run Plays for Each Team", 
       subtitle = "(2022 Season)", x = "", y = "Proportion of Plays",
       fill = "Run or Pass") +
  theme_bw() +
  theme(plot.title = element_text(face="bold", size=16), 
        plot.subtitle = element_text(face="bold"),
        axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(face="bold"),
        axis.text.x = element_text(color = "black", face = "bold"),
        axis.text.y = element_text(color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # Remove panel background
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA) ) +
  theme(axis.text.x = element_text( angle = 32 ) ) +
  annotate( "label", x=28, y=.8, label = "Top: Passes \n Bottom: Runs" )

stack
```

```{r}
ggsave('/Users/garrett/Desktop/stack.png', stack, bg='transparent', 
       width = 21, height = 11, units = "in", scale = .375, dpi = 500 )
```

(Figure 1): The above plot shows the proportion of running and passing plays from each team in the 2022 season. The top grey bar represents the proportion of passing plays. Where as the bottom colored bar represents the proportion of running plays. There are two teams that stand out in this graph, Chicago (CHI) and Tampa Bay (TB). Chicago executes more running plays than every other team. This makes sense as they have quarterback Justin Fields, who is arguably the most mobile QB the league has ever seen. On the other end, we have Tampa Bay executing more passes than runs. This also makes sense as they have quarterback Tom Brady, who is a notorious pocket passer. This visualization is insightful because we can see if teams are leveraging the skills of their players, particularly quarterbacks.



```{r}
stack2 <- ggplot(d, aes(x=possession_abbr, y=count, fill=run_pass ) )+
  geom_col(position = "fill") +
  scale_fill_manual(values = c("grey", d$team_color) ) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Proportion of Pass and Run Plays for Each Team", 
       subtitle = "(2022 Season)", x= "", y= "Proportion of Plays",
       fill = "Run or Pass") +
  theme_bw() +
  theme(plot.title = element_text(face="bold", size=16), 
        plot.subtitle = element_text(face="bold"),
        axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(face="bold"),
        axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # Remove panel background
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA) ) +
  theme(axis.text.x = element_text( angle = 45 ),
        legend.title = element_text(face = "bold"))

#stack2
```


```{r fig.height=5, fig.width=8}
scatter <- ggplot( run_pass_df, aes(x=num_run_per_game, y=num_pass_per_game) ) +
  geom_point(stat = "identity", size=.5, alpha = .001 ) +
  geom_hline(yintercept = mean(run_pass_df$num_pass_per_game), color = "red", linetype = "dashed", alpha=0.8) +
  geom_vline(xintercept =  mean(run_pass_df$num_run_per_game), color = "red", linetype = "dashed", alpha=0.8) +
  geom_nfl_logos(aes(team_abbr = possession_abbr), width = .08, alpha = 0.75)+
  labs(title = "Every Team's Average Number of Passes and Runs per Game", 
       subtitle = "(2022 Season)", x= "Average Number of Runs",
       y= "Average Number of Passes") + 
  theme_bw() +
  theme(plot.title = element_text(face="bold", size=16), 
        plot.subtitle = element_text(face="bold"),
        axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(face="bold"),
        axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # Remove panel background
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA) )
  
scatter
```

```{r}
ggsave('/Users/garrett/Desktop/logo_plot.png', scatter, bg='transparent', 
       width = 21, height = 14, units = "in", scale = .375, dpi = 500 )
```

(Figure 2): The above plot shows every team's average number of passes and runs per game. This plot portrays similar information as the figure 1 but in a different visual way. We can easily compare our favorite teams to on another to see if they execute more running or passing plays. Additional information that this plot provides is the number of plays teams are able to execute in a given game. We can see teams like the Eagles, Giants, and Browns execute an above average number of running plays and an average number of passing plays. Compared to teams like the Bengals and Texan who executed a below average number of running plays and only a average amount of passing plays.















